Designing a high-performance boundary element library with OpenCL and Numba

### 2021-04-06 bem.tex
make ise, isation, etc consistent
add tikz sphere
add command for \ds

The most simple boundary integral equation is of the form
\begin{align}
  \label{eq:bnd_integral}
  \int_{\Gamma} g(\bx, \by)\phi(\by)\,\mathrm{d}s(\by) &= f(\bx),&&\bx\in\Gamma.
\end{align}
The function $g(\bx, \by)$ is a Green's function, $f$ is a given right-hand side, and $\phi$ is an unknown surface density over the boundary $\Gamma$ of a bounded three dimensional domain $\Omega\subset\mathbb{R}^3$.

As a concrete example, we consider computing the electrostatic capacity of an object $\Omega$. In this case, we solve the above equation with $f(\bx)=1$ and $g(\bx,\by)=\frac{1}{4\uppi|\bx - \by|}$. Once $\phi$ has been found, the normalised capacity is then obtained using $c = \frac{1}{4\uppi}\int_{\Gamma}\phi(\bx)\,\mathrm{d}s(\bx)$.

@@

\begin{figure}
  \center
  \includegraphics[width=5cm]{img/sphere}
  \caption{Discretisation of a sphere into flat surface triangles.}
  \label{fig:triangulation}
\end{figure}

@@

To discretise \cref{eq:bnd_integral}, we define a \textbf{test (or dual) space} in terms of a basis \(\psi_1\) to \(\psi_N\). The discrete representation of the above problem then takes the form
$$
Ax = b
$$
with
\begin{align*}
A_{ij} &= \int_{\Gamma}\psi_i(\bx)\int_{\Gamma}g(\bx, \by)\phi_j(\bx)\,\mathrm{d}s(\by)\,\mathrm{d}s(\bx)\\
b_i &= \int_{\Gamma}\psi_i(\bx)f(\bx)\,\mathrm{d}s(\bx).
\end{align*}
In the case of piecewise constant trial and test functions, the definition of $A_{ij}$ simplifies to $A_{ij} = \int_{\Gamma}\int_{\Gamma}g(\bx, \by)\,\mathrm{d}s(\by)\,\mathrm{d}s(\bx)$.

In Bempp-cl, an operator definition consists of the type of the operator (e.g. Laplace single-layer in the above example),
the definition of the trial and test spaces, and the definition of the range space. The range space is required for operator
products and not relevant for the purpose of this paper. The function $f$ is represented as a \textbf{grid function} object, which
consists of either the dual representation in the forms of the integrals $b_i = \int_{\Gamma}\psi_i(\bx)f(\bx)\,\mathrm{d}s(x)$ or directly through its coefficients $f_j$ in the representation $f=\sum_{j=1}^N f_j\phi_j$.

@@

In summary, to solve a boundary integral equation problem, the following steps are performed by Bempp:
\begin{enumerate}
	\item Import of the surface description as triangulation data.
	\item Definition of the trial space, test space and range space.
	\item Discretization into a matrix problem $Ax=b$.
	\item Solution of the matrix problem by either a direct or iterative solver.
	\item Evaluation of domain potential operators for visualization and post-processing.
\end{enumerate}

>>>>

The most simple boundary integral equation is of the form
\begin{align}
  \label{eq:bnd_integral}
  \int_{\Gamma} g(\bx, \by)\phi(\by)\ds[\by] &= f(\bx),&&\bx\in\Gamma.
\end{align}
The function $g(\bx, \by)$ is a Green's function, $f$ is a given right-hand side, and $\phi$ is an unknown surface density over the boundary $\Gamma$ of a bounded three dimensional domain $\Omega\subset\mathbb{R}^3$.

As a concrete example, we consider computing the electrostatic capacity of an object $\Omega$. In this case, we solve the above equation with $f(\bx)=1$ and $g(\bx,\by)=\frac{1}{4\uppi|\bx - \by|}$. Once $\phi$ has been found, the normalised capacity is then obtained using $c = \frac{1}{4\uppi}\int_{\Gamma}\phi(\bx)\ds[\bx]$.

@@

\begin{figure}
  \centering
  \input{img/sphere}
  \caption{Discretisation of a sphere into flat surface triangles.}
  \label{fig:triangulation}
\end{figure}

@@

To discretise \cref{eq:bnd_integral}, we define a \textbf{test (or dual) space} in terms of a basis \(\psi_1\) to \(\psi_N\). The discrete representation of the above problem then takes the form
$$
Ax = b
$$
with
\begin{align*}
A_{ij} &= \int_{\Gamma}\psi_i(\bx)\int_{\Gamma}g(\bx, \by)\phi_j(\bx)\ds[\by]\ds[\bx]\\
b_i &= \int_{\Gamma}\psi_i(\bx)f(\bx)\ds[\bx].
\end{align*}
In the case of piecewise constant trial and test functions, the definition of $A_{ij}$ simplifies to $A_{ij} = \int_{\Gamma}\int_{\Gamma}g(\bx, \by)\ds[\by]\ds[\bx]$.

In Bempp-cl, an operator definition consists of the type of the operator (e.g. Laplace single-layer in the above example),
the definition of the trial and test spaces, and the definition of the range space. The range space is required for operator
products and not relevant for the purpose of this paper. The function $f$ is represented as a \textbf{grid function} object, which
consists of either the dual representation in the forms of the integrals $b_i = \int_{\Gamma}\psi_i(\bx)f(\bx)\ds[\bx]$ or directly through its coefficients $f_j$ in the representation $f=\sum_{j=1}^N f_j\phi_j$.

@@

In summary, to solve a boundary integral equation problem, the following steps are performed by Bempp:
\begin{enumerate}
	\item Import of the surface description as triangulation data.
	\item Definition of the trial space, test space and range space.
	\item Discretisation into a matrix problem $Ax=b$.
	\item Solution of the matrix problem by either a direct or iterative solver.
	\item Evaluation of domain potential operators for visualisation and post-processing.
\end{enumerate}

### 2021-04-06 img/bempp_overview.tex
add command for \ds

\draw[line width=1pt,line join=round,line cap=round,bemppgrey] (0,0) rectangle (4.6,0.5);
\node at (2.3,0.25) {\texttt{bempp.api}};
\draw[line width=1pt,line join=round,line cap=round,bempporange] (0,0.7) rectangle (1,1.2);
\node at (0.5,0.95) {\footnotesize grids};
\draw[line width=1pt,line join=round,line cap=round,bempporange] (1.2,0.7) rectangle (2.2,1.2);
\node at (1.7,0.95) {\footnotesize spaces};
\draw[line width=1pt,line join=round,line cap=round,bempporange] (2.4,0.7) rectangle (3.4,1.2);
\node at (2.9,0.95) [text width=1.45cm,align=center] {\footnotesize grid\\[-1.2mm]functions};
\draw[line width=1pt,line join=round,line cap=round,bempporange] (3.6,0.7) rectangle (4.6,1.2);
\node at (4.1,0.95) {\footnotesize operators};
\draw[line width=1pt,line join=round,line cap=round,bemppgrey] (3.5,1.4) rectangle (4.7,1.9);
\node at (4.1,1.65) {\footnotesize\texttt{bempp.core}};

@@

\draw[line width=1pt,decorate,decoration={brace,amplitude=3pt,mirror}] (3.5,2) -- (4.7,2);
\node[anchor=north] at (4.1,2.05) {\footnotesize Numba / OpenCL};
\end{tikzpicture}

>>>>

\draw[line width=1pt,line join=round,line cap=round,bemppgrey] (0,0) rectangle (4.6,0.5);
\node at (2.3,0.25) {\texttt{bempp.api}};
\draw[line width=1pt,line join=round,line cap=round,bempporange] (0,0.7) rectangle (1,1.2);
\node at (0.5,0.95) {\footnotesize grids};
\draw[line width=1pt,line join=round,line cap=round,bempporange] (1.2,0.7) rectangle (2.2,1.2);
\node at (1.7,0.95) {\footnotesize spaces};
\draw[line width=1pt,line join=round,line cap=round,bempporange] (2.4,0.7) rectangle (3.4,1.2);
\node at (2.9,0.95) [text width=1.45cm,align=center] {\footnotesize grid\\[-1.2mm]functions};
\draw[line width=1pt,line join=round,line cap=round,bempporange] (3.6,0.7) rectangle (4.6,1.2);
\node at (4.1,0.95) {\footnotesize operators};
\draw[line width=1pt,line join=round,line cap=round,bemppgrey] (3.45,1.4) rectangle (4.75,1.9);
\node at (4.1,1.65) {\footnotesize\texttt{bempp.core}};

@@

\draw[line width=1pt,decorate,decoration={brace,amplitude=3pt,mirror}] (3.45,2) -- (4.75,2);
\node[anchor=north] at (4.1,2.05) {\footnotesize Numba / OpenCL};
\end{tikzpicture}
